---
title: "Homework 04"
author: "Qi Yang"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r load, warning = FALSE, message = FALSE}
library(tidyverse)
library(tidyr)
library(gapminder)
library(DT)
guest <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/attend.csv")
email <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/emails.csv")
```

## Exercise 1: Univariate Data Reshaping - Option 1
### 1. Produce a tibble with one row per year, and columns for life expectancy for two or more countries.
```{r}
lifeExp_comp <- gapminder %>% 
  select(year, country, lifeExp) %>%  # Not necessary, saved simply for comparison with data re-lengthening
  filter(country == "Central African Republic" | country == "Japan")
lifeExp_comp_wider <- lifeExp_comp %>%  
                      pivot_wider(id_col = year, 
                      names_from = country,
                      values_from = lifeExp) %>% 
                      rename(lifeExp_CAR = "Central African Republic", lifeExp_Japan = "Japan") %>%  
                      # Rename to make the tibble nicer
                      round(digits = 2) # Round digits to 2
DT::datatable(lifeExp_comp_wider, rownames = FALSE)
```
This tibble provides life expectancies of **Central African Republic (CAR)** and **Japan** in different years. They have the *lowest* and *highest* life expectany respectively in 2019, so I chose them for comparison here. We can see that Japanese people live almost *twice* as long as people in CAR across years.

### 2. Use the new data shape to scatterplot life expectancy for one country against that of another.
```{r}
lifeExp_comp_wider %>% 
  ggplot(aes(lifeExp_CAR, lifeExp_Japan, label = year)) + # Add year labels
  geom_point()+
  labs(x = "Life expectancy of Central African Republic",
       y = "Life expectancy of Japan")+
  geom_text(hjust = 0, nudge_x = -2.1) + # Put year labels at proper places
  theme_classic() +
  coord_fixed() # Make axes on the same scale
```

The scatterplot shows the life expectancy of **Japan** vs. **CAR**. We can see huge differences between their life expectancies more clearly and directly.

### 3. Re-lengthen the data.

```{r}
lifeExp_comp_wider %>% 
   rename("Central African Republic" = "lifeExp_CAR", "Japan" = "lifeExp_Japan") %>% 
   # Give back their original names
   pivot_longer(cols = c(-year),
               names_to = "country",
               values_to = "lifeExp") %>% # Re-lengthening 
   arrange(country) %>% # Arrange rows in the same order as before
   DT::datatable()
```

## Exercise 2: Multivariate Data Reshaping - Option 1
### 1. Make a tibble with one row per year, and columns for life expectancy and GDP per capita for two countries.

```{r}
lifeGDP <- gapminder %>%
           select(year, country, lifeExp, gdpPercap) %>% 
           filter(country == "Central African Republic" | country == "Japan") 
lifeGDP_wider <- lifeGDP %>% 
                 pivot_wider(id_cols     = c(year),
                 names_from  = c(country), 
                 names_sep   = "_", 
                 values_from = c(lifeExp, gdpPercap) ) %>% 
                 round(digits = 2) %>% 
                 rename(lifeExp_CAR = "lifeExp_Central African Republic", gdpPercap_CAR = "gdpPercap_Central African Republic")
DT::datatable(lifeGDP_wider)
```
This tibble shows the life expectancy and GDP per capita of **CAR** and **Japan** across years. We can see tremendous gaps between not only life expectancies but also GDP per capita: e.g. In 1997, Japan's GDP per capita was almost *40* times higher than CAR!

### 2. Re-lengthen the data.

```{r}
lifeGDP_wider %>% 
    rename("lifeExp_Central African Republic" = "lifeExp_CAR", "gdpPercap_Central African Republic" = "gdpPercap_CAR") %>% 
    pivot_longer(cols      = c(-year), 
               names_to  = c(".value", "country"),
               names_sep = "_") %>% 
    arrange(country) # Arrange rows in the same order as before
```


## Exercise 3: Table Joins
### 3.1 Add a column for email address for each guest in the guestlist.

```{r}
email_separate <- email %>% 
                  separate_rows(guest, sep = ", ") %>% # let emails separated for each person
                  rename(name = guest) # rename for table joining
guest %>% 
   left_join(email_separate, by = c("name")) %>% # add email for each person in guest dataset 
   DT::datatable(rownames = FALSE)
```

### 3.2 Show people we have emails for yet are not on the guestlist.

```{r}
email_separate %>% 
  anti_join(guest, by = c("name")) %>% 
  DT::datatable(rownames = FALSE)
```

### 3.3 Guestlist that includes everyone we have emails for.

```{r}
guest %>% 
  full_join(email_separate, by = c("name")) %>% 
  DT::datatable(rownames = FALSE)
```

