---
title: "hw_02_ggplot_gapminder"
author: "Qi Yang"
date: "2019/9/23"
output: 
        html_document:
        keep_md: true
        toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r load, echo = FALSE, warning = FALSE, message = FALSE}
library(tidyverse)
library(gapminder)
```

## Exercise 1: Basic `dplyr`
### 1.1 Subseting gapminder to three countries in the 1970’s with `filter()`.
```{r, warning = FALSE}
gapminder %>%
  filter(year > 1969 & year < 1980,
         country == "Italy" | country == "Australia" | country == "Mexico")
```
### 1.2 Selecting “country” and “gdpPercap” from 1.1 with ` %>% `.
```{r, warning = FALSE}
gapminder %>%
  filter(year > 1969 & year < 1980,
         country == "Italy" | country == "Australia" | country == "Mexico") %>% 
  select(country, gdpPercap)
```
### 1.3 Adding variable `lifeExpInc` to show filtered entries that have experienced drops in life expectancy with `lag()`.
```{r, warning = FALSE}
gapminder %>%
  mutate(lifeExpInc = lifeExp-lag(gapminder$lifeExp,1)) %>% 
  filter(lifeExpInc < 0, year != 1952) %>% 
  select(lifeExpInc, everything()) %>% 
  arrange(lifeExpInc)
```
### 1.4 Adding `gdpPercapMax` to show the max GDP per capita experienced by each country with `max()`.
```{r, warning = FALSE}
gapminder %>% 
  group_by(country) %>% 
  filter(gdpPercap == max(gdpPercap)) %>% 
  rename(gdpPercapMax = gdpPercap) %>% 
  select(gdpPercapMax, country) %>% 
  arrange(desc(gdpPercapMax))
```
### 1.5 Scatter plot: Canada’s life expectancy vs. GDP per capita.
```{r, echo = FALSE, warning = FALSE}
gapminder %>% 
  filter(country == "Canada") %>% {
  ggplot(.) +
  geom_point(aes(gdpPercap, lifeExp)) +
  scale_y_continuous("Life expectancy (year)") +
  scale_x_log10("GDP per capital",
                labels = scales::dollar_format())
  }
```

## Exercise 2: Explore individual variables with `dplyr`
Categorical variable: **country**.  
Quantitative variable: **pop**.  

### 2.1 Possible values/range  
**country**:   
- The number of different countries is:
```{r, warning = FALSE}
gapminder %>% 
  select(country) %>% 
  unique() %>% 
  nrow()
```
So there are 142 different possible values. Since the space is limited, let's just see the first and last few rows of it (arranged alphabetically).   
- The first few values:
```{r, warning = FALSE}
gapminder %>% 
  select(country) %>% 
  unique() %>% 
  head()
```
- The last few values:
```{r, warning = FALSE}
gapminder %>% 
  select(country) %>% 
  unique() %>% 
  tail()
```
**pop**:  
- The range is:
```{r, warning = FALSE}
gapminder %>% 
  select(pop) %>% 
  range()
```
### 2.2 Typical values, spread and distribution
2.2.1 Before we explore the data, we need to know:       
- Measures of typical values: mostly *mean* or *median*.  
- Measures of spread: *range*, *quartiles*, *interquartile range*, *variance* and *standard deviation*.  
- Distribution shows all possible values of data.  

References can be found here:
1. [CSBSJU](http://www.physics.csbsju.edu/stats/descriptive2.html)), 
2. [Australian Bureau of Statistics](https://www.abs.gov.au/websitedbs/a3121120.nsf/home/statistical+language+-+measures+of+spread)
3. [Statistics How to](https://www.statisticshowto.datasciencecentral.com/data-distribution/)  

2.2.2 First let's see typical values and spread.
- The *summary* of pop, which gives us most of the information above.  
```{r, warning = FALSE}
gapminder %>% 
  select(pop) %>% 
  summary(gapminder)
```
- *interquartile range* of pop:  
```{r, warning = FALSE}
  IQR(gapminder$pop)
```
- *variance* of pop:  
```{r, warning = FALSE}
gapminder %>% 
  select(pop) %>% 
  var()
```
(That is HUGE.)  
- *standard deviation* of pop:
```{r}
  sd(gapminder$pop)
```

Let's sum up.  
- Typical values:

| Type  | Value   |
| ------|---------|
| mean  | 29601212|
| median| 7023596 |

- Spread:

| Type  | Value   |
| ------|---------|
| range | lowest: 60011 highest: 1318683096|
| quartiles| Q1:2793664 Q2:7023596 Q3:19585222|
| interquartile range  | 16791558|
| variance| 11269499041091670 |
| standard deviation| 106157897|

2.2.3 Next, let's see the distribution.
**country**:   
```{r}
gapminder %>% 
  select(country) %>% 
  count(country)
```
We can easily tell from the tibble that each country has 12 entries.
**pop**:
We might want to use a graph which is easier to see the distribution.





## Exercise 3: Explore various plot types
### 3.1 Scatterplot: Life expectancy vs. population in Africa and Europe.
```{r, warning = FALSE, echo = FALSE, message = FALSE}
ggplot(subset(gapminder,continent %in% c("Africa" , "Europe")),
       aes(pop/1000000, lifeExp, color=continent)) + 
  geom_point() +
  geom_smooth(se=FALSE, fullrange=TRUE) +
  scale_x_log10("Population (million)") +
  scale_y_continuous("Life expectancy (years)") +
  theme_classic() +
  scale_color_brewer(palette="Paired", name = "Continent")
```  

### 3.2 Bar plot: GDP per capita vs. year in each continent.
```{r, warning = FALSE, echo = FALSE}
ggplot(gapminder, aes(year, gdpPercap/1000)) +
  geom_bar(stat = "identity", 
           aes(fill = continent), 
           position = "dodge") + 
  scale_x_continuous("Year") +
  scale_y_continuous("GDP per capita (thousand)", 
                     labels = scales::dollar_format(), 
                     expand=c(0,0)) +
  theme_classic() + 
  scale_fill_brewer(palette = "Set3", name = "Continent")
```

