---
title: "hw_02"
author: "Qi Yang"
date: "2019/9/23"
output: 
        github_document:
        keep_md: true
        toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r load, echo = FALSE, warning = FALSE, message = FALSE}
options(scipen = 999) # Force R not to use scientific notation
library(tidyverse)
library(gapminder)
```

## Exercise 1: Basic `dplyr`
### 1.1 Subseting gapminder to three countries in the 1970’s with `filter()`.
```{r, warning = FALSE}
gapminder %>%
  filter(year > 1969 & year < 1980,
         country == "Italy" | country == "Australia" | country == "Mexico")
```
### 1.2 Selecting “country” and “gdpPercap” from 1.1 with ` %>% `.
```{r, warning = FALSE}
gapminder %>%
  filter(year > 1969 & year < 1980,
         country == "Italy" | country == "Australia" | country == "Mexico") %>% 
  select(country, gdpPercap)
```
### 1.3 Adding variable `lifeExpInc` to show filtered entries that have experienced drops in life expectancy with `lag()`.
```{r, warning = FALSE}
gapminder %>%
  mutate(lifeExpInc = lifeExp-lag(gapminder$lifeExp,1)) %>% 
  filter(lifeExpInc < 0, year != 1952) %>% 
  select(lifeExpInc, everything()) %>% 
  arrange(lifeExpInc)
```
### 1.4 Adding `gdpPercapMax` to show the max GDP per capita experienced by each country with `max()`.
```{r, warning = FALSE}
gapminder %>% 
  group_by(country) %>% 
  filter(gdpPercap == max(gdpPercap)) %>% 
  rename(gdpPercapMax = gdpPercap) %>% 
  select(gdpPercapMax, country) %>% 
  arrange(desc(gdpPercapMax))
```
### 1.5 Scatter plot: Canada’s life expectancy vs. GDP per capita.
```{r, echo = FALSE, warning = FALSE}
gapminder %>% 
  filter(country == "Canada") %>% {
  ggplot(.) +
  geom_point(aes(gdpPercap, lifeExp)) +
  scale_y_continuous("Life expectancy (year)") +
  scale_x_log10("GDP per capital",
                labels = scales::dollar_format())
  }
```

## Exercise 2: Explore individual variables with `dplyr`  
- Categorical variable: **country**.      
- Quantitative variable: **pop** (population).  

### 2.1 Possible values/range  
**country**:      

- The number of different countries:
```{r, warning = FALSE}
gapminder %>% 
  select(country) %>% 
  unique() %>% 
  nrow()
```
So there are 142 different possible values. Since the space is limited, let's just see the first and last few values (arranged alphabetically). 

- The first few values:  
```{r, warning = FALSE}
gapminder %>% 
  select(country) %>% 
  unique() %>% 
  head()
```
- The last few values:  
```{r, warning = FALSE}
gapminder %>% 
  select(country) %>% 
  unique() %>% 
  tail()
```
**pop**:    

- The range:
```{r, warning = FALSE}
gapminder %>% 
  select(pop) %>% 
  range()
```
### 2.2 Typical values, spread and distribution
#### 2.2.1 Before we explore the data, we need to know:   

- Measures of *typical values*: mostly *mean* or *median*.  

- Measures of *spread*: *range* (see 2.1), *quartiles*, *interquartile range*, *variance* and *standard deviation*.   

(These two types of measures above might not apply to a categorical variable, which in our case is **country**, so we will only talk about these measures of **pop**)

- *Distribution*: all possible values of data.  

References:  
1. [CSBSJU](http://www.physics.csbsju.edu/stats/descriptive2.html)   
2. [Australian Bureau of  Statistics](https://www.abs.gov.au/websitedbs/a3121120.nsf/home/statistical+language+-+measures+of+spread)  
3. [Statistics How to](https://www.statisticshowto.datasciencecentral.com/data-distribution/)  

#### 2.2.2 To start, let's see *typical values* and *spread* of **pop**.  

- Summary (*mean*, *median* and *quartiles*):  
```{r, warning = FALSE}
gapminder %>% 
  select(pop) %>% 
  summary(gapminder)
```
- *Interquartile range*:  
```{r, warning = FALSE}
  IQR(gapminder$pop) # Tried to use piping in dplyr but failed, so I wrote in the orginal way.
```
- *Variance*:  
```{r, warning = FALSE}
gapminder %>% 
  select(pop) %>% 
  var()
```
(That is HUGE)   

- *Standard deviation*:
```{r}
  sd(gapminder$pop) # Tried to use piping in dplyr but failed, so I wrote in the orginal way.
```

- Let's sum up the *typical values* and *spread* of **pop**.   

- Typical values:

| Type  | Value   |
| ------|---------|
| mean  |29601212|
| median|7023596 |

- Spread:

| Type  | Value   |
| ------|---------|
| range | lowest: 60011   highest: 1318683096|
| quartiles| Q1:2793664    Q2:7023596    Q3:19585222|
| interquartile range  | 16791558|
| variance|11269499041091670 |
| standard deviation| 106157897|

#### 2.2.3 Next, let's see the *distribution*.  

**country**:   
```{r}
gapminder %>% 
  select(country) %>% 
  count(country)
```
We can easily tell from the tibble that each country has 12 entries.  

**pop**:   

We might want to use a graph which is easier to see the distribution.  

```{r, warning = FALSE, echo = FALSE}
gapminder %>% 
  select(pop) %>% {
  ggplot(.) + 
      geom_density(aes(x = pop), alpha = .5, fill = "#1c9099") + 
      theme_classic() + 
      scale_x_log10("Population")+
      scale_y_continuous("Density", expand=c(0,0))}

```

## Exercise 3: Explore various plot types
### 3.1 Scatterplot: Life expectancy vs. population in Africa and Europe.
```{r, warning = FALSE, echo = FALSE, message = FALSE}
ggplot(subset(gapminder,continent %in% c("Africa" , "Europe")),
       aes(pop/1000000, lifeExp, color=continent)) + 
  geom_point() +
  geom_smooth(se=FALSE, fullrange=TRUE) +
  scale_x_log10("Population (million)") +
  scale_y_continuous("Life expectancy (years)") +
  theme_classic() +
  scale_color_brewer(palette="Paired", name = "Continent")
```  

### 3.2 Bar plot: GDP per capita vs. year in each continent.
```{r, warning = FALSE, echo = FALSE}
ggplot(gapminder, aes(year, gdpPercap/1000)) +
  geom_bar(stat = "identity", 
           aes(fill = continent), 
           position = "dodge") + 
  scale_x_continuous("Year") +
  scale_y_continuous("GDP per capita (thousand)", 
                     labels = scales::dollar_format(), 
                     expand=c(0,0)) + # make the plot attach to the x-axis
  theme_classic() + 
  scale_fill_brewer(palette = "Set3", name = "Continent")
```

## Recycling
The analyst didn't succeed in getting the entire data for Rwanda and Afghanistan. He or she shouldn't have used `c()` function. A right way to do this job is by running the code below.
```{r}
filter(gapminder, country == "Rwanda" | country == "Afghanistan")
```