---
title: "hw01_gapminder"
author: "Qi Yang"
date: "2019/9/13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Exploring Gapminder
```{r,echo = FALSE}
library(gapminder)
```
1. An **overview** of Gapminder.
```{r,echo = FALSE}
str(gapminder)
```
2. What **variables** does Gapminder have?
```{r,echo = FALSE}
names(gapminder)
```
3. What is the **maximum**, **minimum**, or **mean** value of a variable?  
- Take *population* of all years and of all countries as an example:  
Max: `r max(gapminder$pop)`.   Min: `r min(gapminder$pop)`.   Mean: `r mean(gapminder$pop)`.  
- We can also generate these values at one time. Take *life expectation* of all years and of all countries as an example: 
```{r,echo = FALSE}
summary(gapminder$pop)
```
4. What if we want to know a specific time and country with a certain value (GDP, life expectation, population)?
- The highest GDP per capita in 1972:
```{r, echo = FALSE}
max(gapminder$gdpPercap[which(gapminder$year == "1972")])
```
- Which country has the highest GDP per capita in 1972?
```{r, echo = FALSE}
gapminder$country[which(gapminder$gdpPercap == max(gapminder$gdpPercap[which(gapminder$year == "1972")]))]
```
- The lowest GDP per capita of China:
```{r, echo = FALSE}
min(gapminder$gdpPercap[which(gapminder$country == "China")])
```
- When did China have the lowest GDP?
```{r, echo = FALSE}
gapminder$year[which(gapminder$gdpPercap == min(gapminder$gdpPercap[which(gapminder$country == "China")]))]
```
- If we are interested in a value, we can also point out the year and country of the value. Like this:  
The exact GDP per capita of China in 1972:
```{r, echo = FALSE}
gapminder$gdpPercap[which(gapminder$country == "China" & gapminder$year == "1972")]
```






