---
title: "hw01_gapminder"
author: "Qi Yang"
date: "2019/09/13"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Exploring Gapminder
```{r,echo = FALSE, warning = FALSE}
library(gapminder)
library(tibble)
```
1. An overview of gapminder
```{r,echo = FALSE}
str(gapminder)
```
- How many rows does gapminder have?
```{r,echo = FALSE}
nrow(gapminder)
```
- How many columns does gapminder have?
```{r,echo = FALSE}
ncol(gapminder)
```
- What variables does gapminder have?
```{r,echo = FALSE}
names(gapminder)
```
2. maximum, minimum, mean, range
- What are the maximum, minimum, and mean of *life expectation* of all years and of all countries?
```{r,echo = FALSE}
summary(gapminder$lifeExp)
```
- What is the time range of gapminder?
```{r,echo = FALSE}
range(gapminder$year)
```
3. Find out specific information
- Which country has the highest GDP per capita in 1972?
```{r, echo = FALSE}
gapminder$country[which(gapminder$gdpPercap == max(gapminder$gdpPercap[which(gapminder$year == "1972")]))]
```
- When did Afghanistan have the lowest GDP?
```{r, echo = FALSE}
gapminder$year[which(gapminder$gdpPercap == min(gapminder$gdpPercap[which(gapminder$country == "Afghanistan")]))]
```
4. Extract information based on certain criteria 
- What if we want to view all records about Afghanistan?
```{r, echo = FALSE}
gapminder[gapminder$country == "Afghanistan", ]
```
5. Basic graphic  
- Population changes of all countries over time  
```{r, echo = FALSE}
plot(pop ~ year, gapminder)
```
