---
title: "Homework 05"
author: "Qi Yang"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r load, warning = FALSE, message = FALSE}
library(tidyverse)
library(tidyr)
library(gapminder)
library(DT)
library(forcats)
library(here)
library(gridExtra)
```

## Exercise 1: Explain the value of the here::here package
- One value I found about `here::here` before reading the blog is that it is great for people to run others' codes at their own computer. It doesn't use a path that only exist in the creator's computer. 
- After reading the blog, I found more values about this package.

## Exercise 2: Factor management
### Elaboration for the gapminder data set
1. Drop Oceania.
1) The variables we are going to use are `continent` and `country`. According to the requirement, I need to ensure that they are factors:
```{r}
class(gapminder$continent)
class(gapminder$country)
```
They are indeed factors.     
2) Now I am going to `filter` the gapminder dataset: remove observations associated with the `continent` of Oceania.     

- Before filtering:    
```{r}
nlevels(gapminder$continent)
levels(gapminder$continent)
nrow(gapminder)
```
There are 5 levels in `continent` and 1704 rows.     

- After filtering:     
```{r}
# create new dataset
gap_filtered_oceania <- gapminder %>% 
              filter(continent != "Oceania")
# count levels and rows
nlevels(gap_filtered_oceania$continent)
levels(gap_filtered_oceania$continent)
nrow(gap_filtered_oceania)
```
There are still 5 levels in `continent` but 1680 rows. It means we did filter out some data but the factor Oceania still exist.

3) Now I am going to remove Oceania as an unused factor level.
```{r}
# create new dataset
gap_dropped_oceania <- gap_filtered_oceania %>% 
              droplevels()
# count levels and rows
nlevels(gap_dropped_oceania$continent)
levels(gap_dropped_oceania$continent)
nrow(gap_dropped_oceania)
```
There are only 4 levels in `continent` and 1680 rows. It means we successfully removed some data and the factor Oceania.

2.Reorder the levels of `country`.
In this task, I am going to use the factor `country`. What I aim to do is to see how GDP (= `gdpPercap`*`pop` in `gapminder`) changes over time (the summary statistic) in the top 5 countries which have the highest GDP in 2019, according to [GDP Ranked by Country 2019](http://worldpopulationreview.com/countries/countries-by-gdp/). 

In the following steps, I am going to compare 1) original data, 2) `arrange()`-ed data, and 3) data with releveled factors.

1) Create a new dataset `gapGDP` and view it in tibble and plot.
```{r}
# choose 5 countries
top5gdp <- c("United States", "China", "Japan", "Germany", "India")
# modify gapminder
gapGDP <- gapminder %>% 
  mutate(GDP = gdpPercap*pop/1000000000) %>% # create GDP column
  select(country, GDP, year) %>% # choose variables
  filter(country %in% top5gdp) %>% 
  mutate_if(is.numeric, round) # round the digits of GDP
```
- Original `gapGDP` in tibble:
```{r}
gapGDP %>% 
  DT::datatable()
```
- Original `gapGDP` in plot:
```{r}
ggplot(gapGDP, aes(x = year, y = GDP, # plot before re-leveling
                  color = country)) + 
  geom_line(size = 0.8) +
  labs(x = "Year", y = "GDP (billion dollar)", color = "Country") +
  theme_bw()
```
In both tibble and plot, countries are listed alphabetically: China, Germany, India, Japan, United States.  

2) `gapGDP` after `arrange()` in tibble and plot.

- `arrange()`-ed `gapGDP` in tibble: 
```{r}
gapGDP_arranged <- gapGDP %>% 
  group_by(country) %>% 
  mutate(maxGDP = max(GDP)) %>% 
  arrange(desc(maxGDP)) %>% 
  select(-maxGDP) 
  DT::datatable(gapGDP_arranged)
```

- `arrange()`-ed `gapGDP` in plot:
```{r}
ggplot(gapGDP_arranged,aes(x = year, y = GDP, # plot after arrange()
                  color = country) ) +
  geom_line(size = 0.8) +
  labs(x = "Year", y = "GDP (billion dollar)", color = "Country") +
  theme_bw()
```
We can see that in tibble, countries are arranged in a descending order of their *maximum GDP* (I chose this criteria so that it is consistent with my plot below). However, in plot, countries are not ordered, indicating that `arrange()` can't relevel factors.

3) Re-level factors:
```{r}
ggplot(gapGDP, aes(x = year, y = GDP, # plot after relevling
                  color = fct_reorder2(country, year, GDP, max))) + 
                        # fct_reorder2 is used for 2 variables
  geom_line(size = 0.8) +
  labs(x = "Year", y = "GDP (billion dollar)", color = "Country") +
  theme_bw()
```
After releveling factors, countries are ordered according to their maximum GDP in plot!

## Exercise 3: File input/output (I/O)
1) Create something new.      

In this task, I want to see if life expectancies of each continent in the most recent year (2007) varied. I checked maximum values at first, but there isn't much difference. So let's check out minimum values.
```{r}
gaplifeExp_con <- gapminder %>% 
                  filter(year == "2007") %>% 
                  group_by(continent) %>% 
                  summarise(lifeExpmin = min(lifeExp))
DT::datatable(gaplifeExp_con)
```
2) Export dataset to disk, and reload it back            

- Export:
```{r}
write_csv(gaplifeExp_con, here::here("hw05","gaplifeExp_con.csv"))
```
- Reload:
```{r}
gaplifeExp_con_re <- read_csv(here::here("hw05", "gaplifeExp_con.csv"),
                              col_names = TRUE, 
                              cols(continent = col_factor()))
                              # To make sure `continent` is a factor
DT::datatable(gaplifeExp_con_re)
```
Comment: I didn't have any difficulty writing the data to file. But I had to add lines when reading the exported data back to file, in order to make `continent`, my future "factor", be a real factor. The new file looked just the same as before.    

3) Order my data with my factor        

- Before re-leveling factors:
```{r}
gaplifeExp_con_re %>%
  ggplot() +
  geom_col(aes(x = continent, y = lifeExpmin)) +
  theme_bw() +
  labs(x = "Continent", y = "Minimum life expectancy")
```
The continents are ordered alphabetically.

- After re-leveling factors:
```{r}
gaplifeExp_con_re %>%
  ggplot() +
  geom_col(aes(x = fct_reorder(continent, lifeExpmin), y = lifeExpmin)) +
  theme_bw() +
  labs(x = "Continent", y = "Minimum life expectancy")
```
The continents are plotted in an ascending order of minimum life expectancy. We can see that there indeed were differences. The lowest life expectancy took place in Africa (< 40 years old), and the highest existed in Oceania (> 80 years old).   

## Exercise 4: Visualization design
In this task, I have picked 3 plots I created in my assignment before to modify them.     

1. Changes of life expectancy over time on different continents (Homework03, Task Option 5)
```{r}
hw03_old <- gapminder %>% # old plot
  group_by(continent, year) %>% 
  summarize(min = min(lifeExp),
            mean = mean(lifeExp),
            max = max(lifeExp)) %>% 
  ggplot(aes(year, mean, colour = continent)) +
  geom_line()+
  geom_point()+
  scale_x_continuous("Year")+
  scale_y_continuous("Mean of life expectancy")+
  theme_light()+
  scale_color_brewer(palette="Paired", name = "Continent")

hw03_new <- gapminder %>% # new plot
  group_by(continent, year) %>% 
  summarise(meanlifeExp = mean(lifeExp)) %>% 
ggplot(aes(year, meanlifeExp,
                  color = fct_reorder2(continent, year, meanlifeExp))) + 
  geom_line(size = 0.6) +
  labs(x = "Year", y = "Mean of life expectancy", color = "Continent") +
  theme_classic()
```
```{r, fig.height=5, fig.width=15}
grid.arrange(hw03_old, hw03_new, ncol = 2, nrow = 1)
```
<center> Changes of life expectancy over time on different continents, old(left) and new (right) </center>

|What I changed|Oldl|New|
|----|--------|---|
|factor re-leveling| Orders of continents don't match in lines and legend | Matched!|
|Color| Colors are not very obvious | easy to differentiate from each other|
|Background|Grids in graph are kind of distractive|White, lines are easier to be seen|
|Points|Could be redundant|Removed as the general trend matters most|

2. Changes of GDP per capita in each continent over time (Homework 02, Exercise 2)

```{r}
hw02_old <- ggplot(gapminder, aes(gdpPercap, color = continent)) +
  geom_line(stat = "density", size = 0.8) +
  scale_x_log10("GDP per capita", labels = scales::dollar_format())+
  scale_y_continuous("Density")+
  theme_light()+
  scale_color_brewer(palette="Set1", name = "Continent")

hw02_new <- ggplot(gapminder, aes(gdpPercap, continent) )+
  ggridges::geom_density_ridges()+
  theme_classic()+
  scale_x_continuous("GDP per capita", labels = scales::dollar_format())+
  scale_y_discrete("Continent")
```
```{r, fig.height=5, fig.width=15}
grid.arrange(hw02_old, hw02_new, ncol = 2, nrow = 1)
```
|What I changed|Original|New|
|----|--------|---|
|plot type| Kind of messy and hard to pick a continent out to see its density | Continents' density plots separated and easy to be seen|
|Background|Grids in graph are kind of distractive|White, density plots are easier to be seen|
|Color|Actually it is fine|Since `continent` becomes the y-axis there is no need for colors: things are simpler and help viewers to concentrate on the shape of each plot|


|What I changed|Oldl|New|
|----|--------|---|
|plot type| Kind of messy, hard to pick out a continent's density plot| Continents are separated and easy to be seen|
|Background|Grids in graph are kind of distractive|White, density plots are easier to be seen|
|x-axis scale|`scale_x_log10` might cause misunderstanding|Default `scale_x_continuous`, prevent misunderstanding |
|Color|Actually it is fine | As `continent` becomes the y-axis there is no need for colors: the plot becomes simpler and helps viewers to concentrate on the density shape|


## Exercise 5: Writing figures to file

