---
title: "hw_03"
author: "Qi Yang"
output: 
        html_document:
        css: two_columns.css

---
<style type="text/css">
.twoC {width: 100%}
.clearer {clear: both}
.twoC .table {max-width: 50%; float: right}
.twoC img {max-width: 50%; float: left}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r load, warning = FALSE, message = FALSE}
library(tidyverse)
library(tidyr)
library(gapminder)
library(DT)
library(tsibble)
```

## 1.Task Option 2
#### Requirement: Get the maximum and minimum of GDP per capita for all continents.
1. Tibble  
```{r, warning = FALSE}
gapminder %>% 
  group_by(continent) %>% 
  summarize(Max = max(gdpPercap),
            Country_max = country[gdpPercap == Max],
            Year_max = year[gdpPercap == Max],
            Min = min(gdpPercap),
            Country_min= country[gdpPercap == Min],
            Year_min = year[gdpPercap == Min],) %>% 
  rename(Continent = continent) %>% 
DT::datatable()
```
This tibble shows the **maximum** and **minimum GDP per capita**, corresponding **countries** and **years** in each continent. We can see *Europe* had the maximum GDP per capita of the whole world in 2007, when Norway reached 49357.19017. *Africa* had the minimum GDP per capita across the globe in 2002, when Congo, Dem. Rep. achieved only 241.1658765. 

2. Plot       

```{r}
gapminder %>% 
  group_by(continent) %>% 
  summarize(Max = max(gdpPercap),
            Min = min(gdpPercap),
            Max_forPlot = max(gdpPercap)) %>% # create a column for reordering bars in the plot below
  pivot_longer(cols      = c(Max, Min),
               names_to  = "Max_or_Min",
               values_to = "gdpPercap") %>% 
  ggplot(aes(x= reorder(continent, Max_forPlot), gdpPercap)) +
    geom_bar(stat = "identity", aes(fill = Max_or_Min), position = "dodge")+
     scale_y_log10("GDP per capita", labels = scales::dollar_format())+
     scale_x_discrete("Continent") +
     theme_light()+
     scale_fill_brewer(palette = "Paired", name = "GDP per capita")
```  

This plot shows the **maximum** and **minimum GDP per capit**a of each **continent**. We can easily see that the order of continents with GDP per capita from low to high is: *Africa, Oceania, Americas, Europe, Asia*. It is also interesting to see that *Oceania* tends to have the smallest gap between the highest and the lowest GDP per capita, while *Asia* has the biggest gap.


## 2.Task Option 3
#### Requirement: Look at the spread of GDP per capita within the continents.
1. Tibble
```{r}
gapminder %>% 
  group_by(continent) %>%
  summarize(Range = paste(min(gdpPercap), max(gdpPercap), sep = " ~ "),
            Q1 = quantile(gdpPercap,0.25),
            Q2 = quantile(gdpPercap,0.5),
            Q3 = quantile(gdpPercap,0.75),
            IQR = IQR(gdpPercap),
            VAR = var(gdpPercap),
            SD = sd(gdpPercap)) %>% 
  rename(Continent = continent) %>% 
  DT::datatable()
```
This tibble shows the **spread** of **GDP per capita** within the continents, which includes **range**, **quartiles (Q1, Q2, Q3)**, **interquartile range (IQR)**, **variance (VAR)** and **standard deviation (SD)**. Take *Africa* as an example: it has the smallest IQR and SD, indicating that the GDP per capita of African countries are closer to the mean value than other continents. The tibble is very detailed and accurate, but we still need a plot (below) to get direct information about the distribution of GDP per capita.

2. Plot  

```{r}
ggplot(gapminder, aes(gdpPercap, color = continent)) +
  geom_line(stat = "density", size = 0.8) +
  scale_x_log10("GDP per capita", labels = scales::comma_format())+
  scale_y_continuous("Density")+
  theme_light()+
  scale_color_brewer(palette="Set1", name = "Continent")
```

This plot shows the **distribution** of **GDP per capita** for each continent. We can directly tell that *Oceania* seem to have more "concentrated" and higher values, while the values for *Asia* tend to be lower and spread further. 


## 3.Task Option 5
#### Requirement: How is life expectancy changing over time on different continents?
1. Tibble
```{r}
gapminder %>% 
  group_by(continent,year) %>% 
  summarize(Min = min(lifeExp),
            Mean = mean(lifeExp),
            Max = max(lifeExp)) %>% 
  rename(Continent = continent,
         Year = year) %>% 
  DT::datatable()
  
```
This tibble shows how the **minimum, mean** and **maximum** values of **life expectancy** has changed over **time** on each continent. There is a bunch of data, and we may guess that life expectancy goes up with time generally across the globe. To see the trend more clearly and quickly, we can use the plot below.

2. Plot
```{r}
gapminder %>% 
  group_by(continent, year) %>% 
  summarize(min = min(lifeExp),
            mean = mean(lifeExp),
            max = max(lifeExp)) %>% 
  ggplot(aes(year, mean, colour = continent)) +
  geom_line()+
  geom_point()+
  scale_x_continuous("Year")+
  scale_y_continuous("Mean of life expectancy")+
  theme_light()+
  scale_color_brewer(palette="Paired", name = "Continent")
```

The plot shows how the **mean** value of **life expectancy** of each continent changes over **time**, supporting our guess that people tend to live longer with time going on. However, in the case of *Africa*, life expectancy dropped from 1990 to 2005. We can also see that the growing speed was not the same for different continents and in different times, for instance, the value for *Oceania* grew slowly from 1960 to 1970.

## Optional Exercise
#### Requirement: Get table and figure side-by-side.
My tibbles above seem to be too wide to achieve this task, so I just modified the tibble from Task Option 2 and the corresponding plot as an example).

<div class="twoC">
```{r results = 'asis', echo=FALSE}
gapminder %>% 
  group_by(continent) %>% 
  summarize(Max = max(gdpPercap),
            Year_max = year[gdpPercap == Max],
            Min = min(gdpPercap),
            Year_min = year[gdpPercap == Min],) %>% 
  rename(Continent = continent) %>% 
knitr::kable()
```

```{r, echo = FALSE}
gapminder %>% 
  group_by(continent) %>% 
  summarize(Max = max(gdpPercap),
            Min = min(gdpPercap),
            Max_forPlot = max(gdpPercap)) %>% # create a column for reordering bars in the plot below
  pivot_longer(cols      = c(Max, Min),
               names_to  = "Max_or_Min",
               values_to = "gdpPercap") %>% 
  ggplot(aes(x= reorder(continent, Max_forPlot), gdpPercap)) +
    geom_bar(stat = "identity", aes(fill = Max_or_Min), position = "dodge")+
     scale_y_log10("GDP per capita", labels = scales::dollar_format())+
     scale_x_discrete("Continent") +
     theme_light()+
     scale_fill_brewer(palette = "Paired", name = "GDP per capita")
```  
</div>
<div class="clearer"></div>

why is not it working