---
title: "hw_03"
author: "Qi Yang"
date: "2019/9/30"
output: 
        html_document:
        toc: TRUE
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r load, echo = FALSE, warning = FALSE, message = FALSE}
options(scipen = 999) # Force R not to use scientific notation
library(tidyverse)
library(gapminder)
library(DT)
```

## Task 1: 
### Task Option 2: Get the maximum and minimum of GDP per capita for all continents.
1. Tibble    
This tibble shows...
```{r, warning = FALSE}
gapminder %>% 
  group_by(continent) %>% 
  summarize(max_gdpPercap = max(gdpPercap),
            country_max_gdpPercap = country[gdpPercap == max_gdpPercap],
            year_max_gdpPercap = year[gdpPercap == max_gdpPercap],
            min_gdpPercap = min(gdpPercap),
            country_min_gdpPercap= country[gdpPercap == min_gdpPercap],
            year_min_gdpPercap = year[gdpPercap == min_gdpPercap],) %>% 

DT::datatable()
```
2. Plot
This plot shows...
```{r}
ggplot()
```

## Task 2:
### Task Option 3: Look at the spread of GDP per capita within the continents.
1. Spread should include *range*, *quartiles*, *interquartile range*, *variance* and *standard deviation*. This tibble shows...
```{r}
gapminder %>% 
  group_by(continent) %>%
  summarize(range = paste(min(gdpPercap), max(gdpPercap), sep = " ~ "),
            Q1 = quantile(gdpPercap,0.25),
            Q2 = quantile(gdpPercap,0.5),
            Q3 = quantile(gdpPercap,0.75),
            IQR = IQR(gdpPercap),
            VAR = var(gdpPercap),
            SD = sd(gdpPercap)) %>% 
  DT::datatable()
```
2. Plot  
This plot shows...
```{r}
ggplot()
```
## Task 3:
### Task Option 5: How is life expectancy changing over time on different continents?
1. The tibble shows...
```{r}
gapminder %>% 
  group_by(continent,year) %>% 
  summarize(min = min(lifeExp),
            mean = mean(lifeExp),
            max = max(lifeExp)) %>% 
  DT::datatable()
  
```
2. The plot shows...
```{r}
ggplot()
```


