---
title: "Exploring Gapminder"
author: "Qi Yang"
date: "2019/9/14"
output: 
        github_document:
        keep_md: true
        toc: true
        theme: cerulean
---
```{r setup, include=FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r,echo = FALSE, warning = FALSE}
library(gapminder)
library(tibble)
```
## An overview of gapminder
Let's take a look at some basic information about gapminder.
```{r,echo = FALSE, warning = FALSE}
summary(gapminder)
```
- How many **rows** does gapminder have?
```{r,echo = FALSE, warning = FALSE}
nrow(gapminder)
```
- How many **columns** does gapminder have?
```{r,echo = FALSE, warning = FALSE}
ncol(gapminder)
```
- What **variables* does gapminder have?
```{r,echo = FALSE, warning = FALSE}
names(gapminder)
```
## Maximum, minimum, mean, range
Let's dig deeper into gapminder by finding out the maximum, minimum, mean and range of certain variables. 
- Take **life expectation** of all years and of all countries as an example:
```{r,echo = FALSE, warning = FALSE}
summary(gapminder$lifeExp)
```
- What is the **time** range of gapminder?
```{r,echo = FALSE, warning = FALSE}
range(gapminder$year)
```
## Extraction based on certain criteria 
We can set up a "keyword" to learn everything about it, for instance, a country.  
- What if we want to view all records about **Afghanistan**?
```{r, echo = FALSE, warning = FALSE}
gapminder[gapminder$country == "Afghanistan", ]
```
## Extraction of specific information
We can also view a single piece of information that we want. 
- Which country has **the highest GDP per capita** in **1972**?
```{r, echo = FALSE, warning = FALSE}
gapminder$country[which(gapminder$gdpPercap == max(gapminder$gdpPercap[which(gapminder$year == "1972")]))]
```
- When did **Afghanistan** have **the lowest GDP**?
```{r, echo = FALSE, warning = FALSE}
gapminder$year[which(gapminder$gdpPercap == min(gapminder$gdpPercap[which(gapminder$country == "Afghanistan")]))]
```
## Graphic  
If we want to learn something more directly, we can use graphics based on gapminder. 
- **Population** changes of all countries over time:  
```{r, echo = FALSE, warning = FALSE}
plot(pop ~ year, gapminder)
```